Estimating Predictive Rate-Distortion Curves via Neural Variational Inference

### 2019-06-13 paper.tex
Updates from Overleaf

\subsection{Parameter Estimation and Evaluation}
We optimize the parameters of the neural networks expressing $\phi, \psi, q$ for the objective (\ref{eq:bound}) using Backpropagation and Adam \citep{kingma-adam:-2014}, a standard and widely used gradient descent-based method for optimizing neural networks.
During optimization, we approximate the gradient by taking a single sample from $Z$ (\ref{eq:z-gauss}) per sample trajectory $X_{-M}, \dots, X_{M-1}$ and use the reparametrized gradient estimator introduced by \citet{kingma-auto-encoding-2014}.
This results in an unbiased estimator of the gradient of (\ref{eq:bound}) w.r.t. the parameters of $\phi, \psi, q$.

@@

Given the estimate for prediction loss, we estimate predictiveness $\operatorname{I}[Z, \finitefuture]$ with the following method.
%by estimating $\operatorname{H}[\finitefuture]$ 
Together with $\psi$, we also create an LSTM network that estimates the marginal probability of future observation sequences $P(\finitefuture)$. Let $P_\eta$ be this estimated distribution; we then estimate the predictiveness $\operatorname{I}[Z, \finitefuture] = \operatorname{H}[\finitefuture] - \operatorname{H}[\finitefuture|Z]$ by the difference between the corresponding cross-entropies on the test set~(\cite{mcallester2018formal}):
%$\operatorname{H}[\finitefuture]$ by the cross-entropy of $P_\eta$ on the test data:
\begin{equation}\label{eq:bound-mc-predictiveness}
- \mathbb{E}_{X_{\finitepast\finitefuture}}\left[\log P_\eta(\finitefuture) - \log P_\psi(\finitefuture | Z)\right]
\end{equation}
which we approximate using Monte-Carlo sampling on the test set as in~(\ref{eq:bound-mc-rate}-\ref{eq:bound-mc-distortion}).

@@

>>>>

\subsection{Parameter Estimation and Evaluation}\label{sec:parameter-estimation}
We optimize the parameters of the neural networks expressing $\phi, \psi, q$ for the objective (\ref{eq:bound}) using Backpropagation and Adam \citep{kingma-adam:-2014}, a standard and widely used gradient descent-based method for optimizing neural networks.
During optimization, we approximate the gradient by taking a single sample from $Z$ (\ref{eq:z-gauss}) per sample trajectory $X_{-M}, \dots, X_{M-1}$ and use the reparametrized gradient estimator introduced by \citet{kingma-auto-encoding-2014}.
This results in an unbiased estimator of the gradient of (\ref{eq:bound}) w.r.t. the parameters of $\phi, \psi, q$.

@@

Given the estimate for prediction loss, we estimate predictiveness $\operatorname{I}[Z, \finitefuture]$ with the following method.
%by estimating $\operatorname{H}[\finitefuture]$ 
We use the encoder network that computes the code vector $h$ (\ref{eq:z-gauss}) to also estimate the marginal probability of the past observation sequence, $P_\eta(\finitepast)$. Because we consider stationary processes, we have 
$- \mathbb{E}_{X_{\finitepast\finitefuture}}\left[\log P_\eta(\finitefuture)\right] = - \mathbb{E}_{X_{\finitepast\finitefuture}}\left[\log P_\eta(\finitepast)\right]$.
Using this observation, we estimate the predictiveness $\operatorname{I}[Z, \finitefuture] = \operatorname{H}[\finitefuture] - \operatorname{H}[\finitefuture|Z]$ by the difference between the corresponding cross-entropies on the test set~\cite{mcallester2018formal}:
%$\operatorname{H}[\finitefuture]$ by the cross-entropy of $P_\eta$ on the test data:
\begin{equation}\label{eq:bound-mc-predictiveness}
- \mathbb{E}_{X_{\finitepast\finitefuture}}\left[\log P_\eta(\finitefuture) - \log P_\psi(\finitefuture | Z)\right],
\end{equation}
which we approximate using Monte-Carlo sampling on the test set as in~(\ref{eq:bound-mc-rate}-\ref{eq:bound-mc-distortion}).

During optimization, we add the cross-entropy term $- \mathbb{E}_{X_{\finitepast\finitefuture}}\left[\log P_\eta(\finitepast)\right]$ to the PRD objective~(\ref{eq:bound}), so that the full training objective comes out to:
\begin{equation}\label{eq:bound-augmented}
	\min_{\phi, \psi, q,\eta} \left[\mathbb{E}_{\finitepast, \finitefuture}	\log P_\eta(\finitepast) + \left[-	\mathbb{E}_{Z \sim \phi(\finitepast)}\left[\log P_\psi(\finitefuture | Z)\right] + \lambda \cdot \operatorname{D_{KL}} [ \infdivx{P_\phi(Z|\finitepast)}{q(Z)}\right]\right].
\end{equation} % should be P_\psi (X_{0...M}|z)? % done (mhahn)
Again by Gibbs' inequality and Propositions~\ref{prop:mi-bound} and~\ref{prop:kl-bound}, this is minimized when $P_\eta$ represents the true distribution over length-$M$ blocks $P(\finitepast)$, $P_\phi(Z|\finitepast)$ describes an optimal code for the given $\lambda$, $q$ is its marginal distribution, and $P_\psi(\finitefuture|Z)$ is the Bayes-optimal decoder.
For approximate solutions to this augmented objective, the inequalities~(\ref{eq:bound1-new}-\ref{eq:bound2-new}) will also remain true due to Propositions~\ref{prop:mi-bound} and~\ref{prop:kl-bound}.

@@

Training used Early Stopping using the development set, as described in Section~\ref{sec:parameter-estimation}.
Models were trained on a TITAN Xp graphics card.
